### 2025-03-26 ms/ms.tex
clean

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[default]{sourcesanspro}
\usepackage{sourcecodepro}
\usepackage{lineno}
\usepackage{setspace}
\doublespacing
\linenumbers
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Fig.}
\else
  \newcommand\figurename{Fig.}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

\hypertarget{abstract}{%
\section*{Abstract}\label{abstract}}
\addcontentsline{toc}{section}{Abstract}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Mapping terrestrial carbon stocks is crucial to successfully
implementing climate change mitigation programs
(\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}). The accuracy of
carbon stock estimates depends on the availability of reliable
allometric models to deduce the aboveground biomass (AGB) of trees from
census data (\protect\hyperlink{ref-Chave2014}{Chave et al. 2014}).
Selecting an appropriate allometric model is, therefore, a critical step
in estimating forest carbon biomass (e.g.,
\protect\hyperlink{ref-Chave2014}{Chave et al. 2014};
\protect\hyperlink{ref-Molto2014}{Molto et al. 2014}). Tree allometry
describes the correlations between different biometric measurements of
trees, such as diameter at breast height (DBH), tree height, and tree
crown size (e.g., \protect\hyperlink{ref-Shinozaki1964a}{Shinozaki et
al. 1964}; \protect\hyperlink{ref-Niklas1994}{Niklas 1994};
\protect\hyperlink{ref-West1999}{West et al. 1999}). These relationships
are widely utilized to estimate tree architecture from DBH and to
estimate forest biomass and production, serving as fundamental
components of dynamic vegetation models (e.g.,
\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
\protect\hyperlink{ref-Falster2017}{Falster et al. 2017};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}). Despite significant
progress, developing precise allometric models across regions
characterized by diverse tree species composition and varying forest
structures remains a persistent challenge. This complexity highlights
the need for improved global-scale allometric models that capture the
diversity of tree forms and ecological contexts.

Identifying the most appropriate mathematical representation of
allometric relationships is a subject of ongoing debate. While numerous
papers have used various functional forms for allometric equations
(e.g., \protect\hyperlink{ref-Thomas1996}{Thomas 1996};
\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}), power-law functions
remain the most common choice for describing the allometric scaling of
tree size variables, including height and crown size with DBH (e.g.,
\protect\hyperlink{ref-Sileshi2023}{Sileshi et al. 2023};
\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}). The
relationships among tree size variables often involve complex
interactions influenced by various factors, such as species diversity,
environmental conditions, and individual variability (e.g.,
\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a};
\protect\hyperlink{ref-Laurans2024}{Laurans et al. 2024}). Consequently,
power-law models often fail to accurately capture the allometric
relationships across the full range of tree sizes, leading to
underestimation of dimensions for smaller trees and overestimation for
larger trees (e.g., \protect\hyperlink{ref-Fayolle2016}{Fayolle et al.
2016}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). This
highlights the need to adopt alternative functional forms---such as
saturating relationships---that can more accurately reflect the
complexity of allometric scaling in vegetation models (e.g.,
\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}).

Earth observation (EO) technologies, including remote sensing and Light
Detection and Ranging (LiDAR), have revolutionized the monitoring of
forests (e.g., \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017})
by enabling the identification and measurement of individual tree
heights and crown dimensions using airborne imagery (e.g.,
\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). However,
because existing allometric models still rely on stem diameter as a
crucial input for biomass estimation, this reliance presents a challenge
for effectively integrating EO technologies into AGB estimation
(\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). While
ground-based LiDAR can measure DBH directly and is much more efficient
than traditional ground surveys
(\protect\hyperlink{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is still
less common and convenient than aerial LiDAR. Alternative approaches
focusing on tree height and crown dimensions as model predictors are
also necessary for remote sensing applications
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). One potential
solution is to estimate DBH from tree height and crown size and then use
established biomass equations, which could be an effective strategy for
integrating airborne imagery into carbon monitoring programs
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).

Accurately representing tree allometry relationships also requires
considering underlying ecological constraints, particularly the
interplay of plant functional traits and their trade-offs influence
individual tree architecture and the composition of entire forest
communities (e.g., \protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). Under
closed-canopy conditions, strong light competition makes the balance
between vertical and horizontal space use crucial for species
coexistence (e.g., \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka
2009}; \protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Among
functional traits, wood density is closely linked to both plant life
history strategies (\protect\hyperlink{ref-Fajardo2022a}{Fajardo 2022})
and light demand (e.g., \protect\hyperlink{ref-King2006}{King et al.
2006}; \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a}). Several
studies suggest that wood density explains a trade-off between rapid
stem expansion and broader crowns, shaping mechanical constraints on
species with diverse architecture and light capture strategies
(\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). However, these
types of studies have examined this relationship post hoc, relating
species-specific allometric parameters to wood density and risking Type
I or Type II errors if the hierarchical data structure is overlooked
(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}). One
exception is Martínez Cano et al.
(\protect\hyperlink{ref-MartinezCano2019}{2019}), who incorporated wood
density into a hierarchical model of tropical trees and concluded it
offered minimal explanatory power for interspecific variation.
Therefore, the role of wood density in global allometric scaling remains
insufficiently explored.

In this study, we present a robust Bayesian hierarchical approach to
delineate allometric relationships involving DBH, tree height (H), crown
radius (CR), and their interspecific variations. Using large datasets
from the Tallo and TRY databases
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020};
\protect\hyperlink{ref-Jucker2022}{Jucker et al. 2022}), which span a
diverse array of tree forms across global forest types and climatic
conditions, we overcome the limitations of previous studies that often
relied on site-specific or regional observations with limited samples
and species. By integrating species-level wood density (WD) into a
multilevel model, our methodology avoids the post-hoc correlations used
in previous studies, offering a more universal and scalable model,
effectively capturing variability from the species to the community
level. Our main objective is to develop models informed by tree
allometric variables (H, CR, DBH, WD) to capture interspecific variation
in allometric scaling globally and offer an improved framework for
estimating above ground biomass. Specifically, we ask the following
questions: (1) What is the best function to estimate tree architecture
from DBH and \emph{vice versa}? (2) How does wood density shape
interspecific variation in these allometric scaling relationships? (3)
How does the choice of allometric relationships affect subsequent AGB
estimation?

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

Our analysis combined tree allometry data from Tallo, which includes
standardized measurements of DBH, height, and crown radius for 5,163
species across various biomes (\protect\hyperlink{ref-Jucker2022}{Jucker
et al. 2022}, https://doi.org/10.5281/zenodo.6637599) along with wood
density data for 10,995 species from TRY version 6
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020},
https://www.try-db.org/TryWeb/Home.php) (see Supplementary Table S1 for
further details). We thoroughly processed the Tallo dataset by removing
998 outliers following Jucker et al.
(\protect\hyperlink{ref-Jucker2022}{2022}). To ensure consistency, we
selected only species with standardized wood density values in g
cm\textsuperscript{-3} from TRY and removed any records with missing
data. Subsequently, we created an overlapping dataset between Tallo and
TRY, resulting in a combined dataset encompassing allometric and wood
density information for 355,881 individuals across 2,961 species.

To reduce computational complexity and ensure the robustness of the
models, we further filtered species with fewer than 20 individuals and
randomly sampled up to 100 individuals per species. This step was
necessary because the dataset is unbalanced; for example, \emph{Quercus
ilex} alone accounts for over 20,000 individuals, representing
approximately 10\% of the dataset. However, \emph{Quercus ilex} is not
necessarily the most abundant species globally. Its prevalence in this
dataset is a result of sampling and does not reflect its true global
abundance. By limiting the number of individuals per species, we
minimize bias due to data imbalance and prevent the overrepresentation
of any single abundant species
(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}), thereby
ensuring that the model remains broadly applicable. Since not all trees
have a complete set of information, we created specific sub-datasets to
select the best predictive model for each allometric relationship for
two clades (i.e., angiosperms and gymnosperms; see below and
Supplementary Table S2 for more details).

\hypertarget{tree-allometry-models}{%
\subsection{Tree allometry models}\label{tree-allometry-models}}

@@

\begin{equation}\protect\hypertarget{eq-like}{}{
\log(y_{i,j}) \sim N(f(X_{i,j}, \pmb{\beta}_{j}), \sigma_y),
}\label{eq-like}\end{equation}

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[default]{sourcesanspro}
\usepackage{sourcecodepro}
\usepackage{lineno}
\usepackage{setspace}
\doublespacing
\linenumbers
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Fig.}
\else
  \newcommand\figurename{Fig.}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}

@@

\section*{Abstract}\label{abstract}
\addcontentsline{toc}{section}{Abstract}

@@

\section{Introduction}\label{introduction}

Mapping terrestrial carbon stocks is crucial to successfully
implementing climate change mitigation programs
(\citeproc{ref-Chave2014}{Chave et al. 2014}). The accuracy of carbon
stock estimates depends on the availability of reliable allometric
models to deduce the aboveground biomass (AGB) of trees from census data
(\citeproc{ref-Chave2014}{Chave et al. 2014}). Selecting an appropriate
allometric model is, therefore, a critical step in estimating forest
carbon biomass (e.g., \citeproc{ref-Chave2014}{Chave et al. 2014};
\citeproc{ref-Molto2014}{Molto et al. 2014}). Tree allometry describes
the correlations between different biometric measurements of trees, such
as diameter at breast height (DBH), tree height, and tree crown size
(e.g., \citeproc{ref-Shinozaki1964a}{Shinozaki et al. 1964};
\citeproc{ref-Niklas1994}{Niklas 1994}; \citeproc{ref-West1999}{West et
al. 1999}). These relationships are widely utilized to estimate tree
architecture from DBH and to estimate forest biomass and production,
serving as fundamental components of dynamic vegetation models (e.g.,
\citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
\citeproc{ref-Falster2017}{Falster et al. 2017};
\citeproc{ref-Song2024}{Song et al. 2024}). Despite significant
progress, developing precise allometric models across regions
characterized by diverse tree species composition and varying forest
structures remains a persistent challenge. This complexity highlights
the need for improved global-scale allometric models that capture the
diversity of tree forms and ecological contexts.

Identifying the most appropriate mathematical representation of
allometric relationships is a subject of ongoing debate. While numerous
papers have used various functional forms for allometric equations
(e.g., \citeproc{ref-Thomas1996}{Thomas 1996};
\citeproc{ref-Feldpausch2011}{Feldpausch et al. 2011};
\citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019};
\citeproc{ref-Song2024}{Song et al. 2024}), power-law functions remain
the most common choice for describing the allometric scaling of tree
size variables, including height and crown size with DBH (e.g.,
\citeproc{ref-Sileshi2023}{Sileshi et al. 2023};
\citeproc{ref-Jucker2024}{Jucker et al. 2024}). The relationships among
tree size variables often involve complex interactions influenced by
various factors, such as species diversity, environmental conditions,
and individual variability (e.g., \citeproc{ref-Poorter2012a}{Poorter et
al. 2012a}; \citeproc{ref-Laurans2024}{Laurans et al. 2024}).
Consequently, power-law models often fail to accurately capture the
allometric relationships across the full range of tree sizes, leading to
underestimation of dimensions for smaller trees and overestimation for
larger trees (e.g., \citeproc{ref-Fayolle2016}{Fayolle et al. 2016};
\citeproc{ref-Song2024}{Song et al. 2024}). This highlights the need to
adopt alternative functional forms---such as saturating
relationships---that can more accurately reflect the complexity of
allometric scaling in vegetation models (e.g.,
\citeproc{ref-Goodman2014}{Goodman et al. 2014};
\citeproc{ref-Song2024}{Song et al. 2024}).

Earth observation (EO) technologies, including remote sensing and Light
Detection and Ranging (LiDAR), have revolutionized the monitoring of
forests (e.g., \citeproc{ref-Jucker2017}{Jucker et al. 2017}) by
enabling the identification and measurement of individual tree heights
and crown dimensions using airborne imagery (e.g.,
\citeproc{ref-Shendryk2016}{Shendryk et al. 2016};
\citeproc{ref-Jucker2017}{Jucker et al. 2017}). However, because
existing allometric models still rely on stem diameter as a crucial
input for biomass estimation, this reliance presents a challenge for
effectively integrating EO technologies into AGB estimation
(\citeproc{ref-Shendryk2016}{Shendryk et al. 2016};
\citeproc{ref-Jucker2017}{Jucker et al. 2017}). While ground-based LiDAR
can measure DBH directly and is much more efficient than traditional
ground surveys (\citeproc{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is
still less common and convenient than aerial LiDAR. Alternative
approaches focusing on tree height and crown dimensions as model
predictors are also necessary for remote sensing applications
(\citeproc{ref-Jucker2017}{Jucker et al. 2017}). One potential solution
is to estimate DBH from tree height and crown size and then use
established biomass equations, which could be an effective strategy for
integrating airborne imagery into carbon monitoring programs
(\citeproc{ref-Jucker2017}{Jucker et al. 2017}).

Accurately representing tree allometry relationships also requires
considering underlying ecological constraints, particularly the
interplay of plant functional traits and their trade-offs influence
individual tree architecture and the composition of entire forest
communities (e.g., \citeproc{ref-Iida2012}{Iida et al. 2012};
\citeproc{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
\citeproc{ref-Yang2023}{Yang and Swenson 2023}). Under closed-canopy
conditions, strong light competition makes the balance between vertical
and horizontal space use crucial for species coexistence (e.g.,
\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\citeproc{ref-Iida2012}{Iida et al. 2012}). Among functional traits,
wood density is closely linked to both plant life history strategies
(\citeproc{ref-Fajardo2022a}{Fajardo 2022}) and light demand (e.g.,
\citeproc{ref-King2006}{King et al. 2006}; \citeproc{ref-Aiba2009}{Aiba
and Nakashizuka 2009}; \citeproc{ref-Poorter2012a}{Poorter et al.
2012a}). Several studies suggest that wood density explains a trade-off
between rapid stem expansion and broader crowns, shaping mechanical
constraints on species with diverse architecture and light capture
strategies (\citeproc{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\citeproc{ref-Iida2012}{Iida et al. 2012}; \citeproc{ref-Yang2023}{Yang
and Swenson 2023}). However, these types of studies have examined this
relationship post hoc, relating species-specific allometric parameters
to wood density and risking Type I or Type II errors if the hierarchical
data structure is overlooked (\citeproc{ref-Gelman2006}{Gelman and Hill
2006}). One exception is Martínez Cano et al.
(\citeproc{ref-MartinezCano2019}{2019}), who incorporated wood density
into a hierarchical model of tropical trees and concluded it offered
minimal explanatory power for interspecific variation. Therefore, the
role of wood density in global allometric scaling remains insufficiently
explored.

In this study, we present a robust Bayesian hierarchical approach to
delineate allometric relationships involving DBH, tree height (H), crown
radius (CR), and their interspecific variations. Using large datasets
from the Tallo and TRY databases (\citeproc{ref-Kattge2020}{Kattge et
al. 2020}; \citeproc{ref-Jucker2022}{Jucker et al. 2022}), which span a
diverse array of tree forms across global forest types and climatic
conditions, we overcome the limitations of previous studies that often
relied on site-specific or regional observations with limited samples
and species. By integrating species-level wood density (WD) into a
multilevel model, our methodology avoids the post-hoc correlations used
in previous studies, offering a more universal and scalable model,
effectively capturing variability from the species to the community
level. Our main objective is to develop models informed by tree
allometric variables (H, CR, DBH, WD) to capture interspecific variation
in allometric scaling globally and offer an improved framework for
estimating above ground biomass. Specifically, we ask the following
questions: (1) What is the best function to estimate tree architecture
from DBH and \emph{vice versa}? (2) How does wood density shape
interspecific variation in these allometric scaling relationships? (3)
How does the choice of allometric relationships affect subsequent AGB
estimation?

\section{Materials and Methods}\label{materials-and-methods}

\subsection{Data}\label{data}

Our analysis combined tree allometry data from Tallo, which includes
standardized measurements of DBH, height, and crown radius for 5,163
species across various biomes (\citeproc{ref-Jucker2022}{Jucker et al.
2022}, https://doi.org/10.5281/zenodo.6637599) along with wood density
data for 10,995 species from TRY version 6
(\citeproc{ref-Kattge2020}{Kattge et al. 2020},
https://www.try-db.org/TryWeb/Home.php) (see Supplementary Table S1 for
further details). We thoroughly processed the Tallo dataset by removing
998 outliers following Jucker et al. (\citeproc{ref-Jucker2022}{2022}).
To ensure consistency, we selected only species with standardized wood
density values in g cm\textsuperscript{-3} from TRY and removed any
records with missing data. Subsequently, we created an overlapping
dataset between Tallo and TRY, resulting in a combined dataset
encompassing allometric and wood density information for 355,881
individuals across 2,961 species.

To reduce computational complexity and ensure the robustness of the
models, we further filtered species with fewer than 20 individuals and
randomly sampled up to 100 individuals per species. This step was
necessary because the dataset is unbalanced; for example, \emph{Quercus
ilex} alone accounts for over 20,000 individuals, representing
approximately 10\% of the dataset. However, \emph{Quercus ilex} is not
necessarily the most abundant species globally. Its prevalence in this
dataset is a result of sampling and does not reflect its true global
abundance. By limiting the number of individuals per species, we
minimize bias due to data imbalance and prevent the overrepresentation
of any single abundant species (\citeproc{ref-Gelman2006}{Gelman and
Hill 2006}), thereby ensuring that the model remains broadly applicable.
Since not all trees have a complete set of information, we created
specific sub-datasets to select the best predictive model for each
allometric relationship for two clades (i.e., angiosperms and
gymnosperms; see below and Supplementary Table S2 for more details).

\subsection{Tree allometry models}\label{tree-allometry-models}

@@

\begin{equation}\phantomsection\label{eq-like}{
\log(y_{i,j}) \sim N(f(X_{i,j}, \pmb{\beta}_{j}), \sigma_y),
}\end{equation}

where \(f(\cdot)\) predicts the expected natural log tree height or
crown radius from observed DBH (and \emph{vice versa}),
\(\pmb{\beta}_{j}=\{a_j, b_j, k_j\}\) denotes the vector of
species-specific parameters, and \(\sigma_y\) represents the standard
deviation of the observation error. We considered three functional forms
for the relationship between tree allometric variables
(\citeproc{ref-Sullivan2018}{Sullivan et al. 2018};
\citeproc{ref-MartinezCano2019}{Martínez Cano et al. 2019}):

Power-law: \begin{equation}\phantomsection\label{eq-pl}{
y = aX^b,
}\end{equation}

Generalized Michaelis-Menten (gMM):
\begin{equation}\phantomsection\label{eq-gMM}{
y = \frac{aX^b}{k + X^b},
}\end{equation}
